---
pagetitle: Leaflet for R - Markers
---

## Markers

Use markers to call out points on the map. Marker locations are expressed in latitude/longitude coordinates, and can either appear as icons or as circles.

### Data sources

Point data for markers can come from a variety of sources:

- `SpatialPoints` or `SpatialPointsDataFrame` objects (from the `sp` package)
- `POINT` objects (from the `sf` package); only `X` and `Y` dimensions will be considered
- Two-column numeric matrices (first column is longitude, second is latitude)
- Data frame with latitude and logitude columns. You can explicitly tell the marker function which columns contain the coordinate data (e.g. `addMarkers(lng = ~Longitude, lat = ~Latitude)`), or let the function look for columns named `lat`/`latitude` and `lon`/`lng`/`long`/`longitude` (case insensitive).
- Simply provide numeric vectors as `lng` and `lat` arguments

Note that `MULTIPOINT` objects from `sf` are not supported at this time.

### Icon Markers

Icon markers are added using the `addMarkers` or the `addAwesomeMarkers` functions. Their default appearance is a dropped pin. As with most layer functions, the `popup` argument can be used to add a message to be displayed on click, and the `label` option can be used to display a text label either on hover or statically.

```{r}
data(quakes)

# Show first 20 rows from the `quakes` dataset
leaflet(data = quakes[1:20,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, popup = ~as.character(mag), label = ~as.character(mag))
```

#### Customizing Marker Icons

You can provide custom markers in one of several ways, depending on the scenario. For each of these ways, the icon can be provided as either a URL or as a file path.

For the simple case of applying a single icon to a set of markers, use `makeIcon()`.

```{r fig.height=1.75}
greenLeafIcon <- makeIcon(
  iconUrl = "http://leafletjs.com/examples/custom-icons/leaf-green.png",
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes[1:4,]) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = greenLeafIcon)
```

If you have several icons to apply that vary only by a couple of parameters (i.e. they share the same size and anchor points but have different URLs), use the `icons()` function. `icons()` performs similarly to `data.frame()`, in that any arguments that are shorter than the number of markers will be recycled to fit.

```{r fig.height=2}
quakes1 <- quakes[1:10,]

leafIcons <- icons(
  iconUrl = ifelse(quakes1$mag < 4.6,
    "http://leafletjs.com/examples/custom-icons/leaf-green.png",
    "http://leafletjs.com/examples/custom-icons/leaf-red.png"
  ),
  iconWidth = 38, iconHeight = 95,
  iconAnchorX = 22, iconAnchorY = 94,
  shadowUrl = "http://leafletjs.com/examples/custom-icons/leaf-shadow.png",
  shadowWidth = 50, shadowHeight = 64,
  shadowAnchorX = 4, shadowAnchorY = 62
)

leaflet(data = quakes1) %>% addTiles() %>%
  addMarkers(~long, ~lat, icon = leafIcons)
```

Finally, if you have a set of icons that vary in multiple parameters, it may be more convenient to use the `iconList()` function. It lets you create a list of (named or unnamed) `makeIcon()` icons, and select from that list by position or name.

```{r fig.height=1.75}
# Make a list of icons. We'll index into it based on name.
oceanIcons <- iconList(
  ship = makeIcon("ferry-18.png", "ferry-18@2x.png", 18, 18),
  pirate = makeIcon("danger-24.png", "danger-24@2x.png", 24, 24)
)

# Some fake data
df <- sp::SpatialPointsDataFrame(
  cbind(
    (runif(20) - .5) * 10 - 90.620130,  # lng
    (runif(20) - .5) * 3.8 + 25.638077  # lat
  ),
  data.frame(type = factor(
    ifelse(runif(20) > 0.75, "pirate", "ship"),
    c("ship", "pirate")
  ))
)

leaflet(df) %>% addTiles() %>%
  # Select from oceanIcons based on df$type
  addMarkers(icon = ~oceanIcons[type])
```

#### Awesome Icons

Leaflet supports even more customizable markers using the [awesome markers](https://github.com/lvoogdt/Leaflet.awesome-markers) leaflet plugin.


The `addAwesomeMarkers()` function is similar to `addMarkers()` function but additionally allows you to specify custom colors for the markers as well as icons from the [Font Awesome](http://fontawesome.io/icons/), [Bootstrap Glyphicons](https://getbootstrap.com/components/), and [Ion icons](http://ionicons.com/) icon libraries.

Similar to the `makeIcon`, `icons`, and `iconList` functions described above, you have `makeAwesomeIcon`, `awesomeIcons` and `awesomeIconList` functions, which enable you to add awesome icons.

```{r}
# first 20 quakes
df.20 <- quakes[1:20,]

getColor <- function(quakes) {
  sapply(quakes$mag, function(mag) {
  if(mag <= 4) {
    "green"
  } else if(mag <= 5) {
    "orange"
  } else {
    "red"
  } })
}

icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(df.20)
)

leaflet(df.20) %>% addTiles() %>%
  addAwesomeMarkers(~long, ~lat, icon=icons, label=~as.character(mag))

```

The `library` argument has to be one of 'ion', 'fa', or 'glyphicon'. The `icon` argument needs to be the name of any valid icon supported by the the respective library (w/o the prefix of the library name).

#### Marker Clusters

When there are a large number of markers on a map, you can cluster them using the [Leaflet.markercluster](https://github.com/Leaflet/Leaflet.markercluster) plug-in. To enable this plug-in, you can provide a list of options to the argument `clusterOptions`, e.g.

```{r fig.height=2.5, message=FALSE}
leaflet(quakes) %>% addTiles() %>% addMarkers(
  clusterOptions = markerClusterOptions()
)
```

Using the `freezeAtZoom` argument of the `markerClusterOptions()` function you can set the clustering to freeze as a specific zoom level. For example `markerClusterOptions(freezeAtZoom = 5)` will freeze the cluster at zoom level 5 regardless of the user's actual zoom level.

### Circle Markers

Circle markers are much like regular circles (see [Lines and Shapes](shapes.html)), except that their radius in onscreen pixels stays constant regardless of zoom level.

You can use their default appearance:

```{r fig.height=1.75}
leaflet(df) %>% addTiles() %>% addCircleMarkers()
```

Or customize their color, radius, stroke, opacity, etc.

```{r fig.height=1.75}
# Create a palette that maps factor levels to colors
pal <- colorFactor(c("navy", "red"), domain = c("ship", "pirate"))

leaflet(df) %>% addTiles() %>%
  addCircleMarkers(
    radius = ~ifelse(type == "ship", 6, 10),
    color = ~pal(type),
    stroke = FALSE, fillOpacity = 0.5
  )
```
