---
title: "An Introduction to the leaflet Package"
author: "Joe Cheng and Yihui Xie"
date: "`r Sys.Date()`"
output:
  knitr:::html_vignette:
    toc: yes
    number_sections: yes
  html_document:
    toc: yes
    number_sections: yes
    theme: cosmo
    css: styles.css
    self_contained: no
    includes:
      before_body: before_body.html
      after_body: after_body.html
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{An Introduction to the leaflet Package}
---
<div class="leaf-separator hide"><span class="glyphicon glyphicon-leaf"></span></div>
<div class="mainbody">

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE)
set.seed(109)
```
[Leaflet](http://leafletjs.com) is an open-source JavaScript library for interactive maps. This R package makes it easy to integrate and control Leaflet maps in R.

# Installation and Use

To install this R package, run this command at your R prompt:

```{r eval=FALSE}
devtools::install_github("rstudio/leaflet")
```

Once installed, you can use this package at the R console, within [R Markdown](http://rmarkdown.rstudio.com/) documents, and within [Shiny](http://shiny.rstudio.com/) applications.

To create a Leaflet map, first you call the function `leaflet()`, then you can add more layers on the map, and manipulate the view. For example,

```{r}
library(leaflet)
m = leaflet() %>% addTiles()
m  # a map with the default OSM tile layer

m = m %>% setView(-93.65, 42.0285, zoom = 17)
m

m %>% addPopups(-93.65, 42.0285, 'Here is the <b>Department of Statistics</b>, ISU')
```

Although we have tried to provide an R-like interface to Leaflet, you may want to check out the [API documentation](http://leafletjs.com/reference.html) of Leaflet occasionally when the meanings of certain parameters are not clear to you.

# The Map Widget

The function `leaflet()` returns a Leaflet map widget, which stores a list of objects that can be modified or updated later. Most functions in this package have an argument `map` as their first argument, which makes it easy to use the pipe operator `%>%` in the **magarittr** package, as you have seen from the examples above.

## Map Methods

You can manipulate the attributes of the map widget using a series of methods. Please see the help page `?setView` for details.

- `setView()` sets the center of the map view and the zoom level;
- `fitBounds()` fits the view into the rectangle `[lng1, lat1]` -- `[lng2, lat2]`;
- `clearBounds()` clears the bound, so that the view will be automatically determined by the range of latitude/longitude data in the map layers if provided;

## The Data Object

Both `leaflet()` and the map layer functions have an argument `data`, and we can reference or derive variables from this data object. Currently these objects are supported:

- From base R:
    - lng/lat matrix
    - data frame with lng/lat columns
- From the [**sp** package](http://cran.rstudio.com/package=sp):
    - SpatialPoints[DataFrame]
    - Line/Lines
    - SpatialLines[DataFrame]
    - Polygon/Polygons
    - SpatialPolygons(DataFrame)
- From the [**maps** package](http://cran.rstudio.com/package=maps):
    - the data frame from returned from `map()`

Perhaps the most frequently used variables in the data objects are the latitudes and longitudes of graphical elements. It is straightforward to derive these variables from **sp** objects since they are well defined in the S4 objects. For a normal matrix or data frame, these two variables are guessed from the column names:

- the latitude variable is guessed by matching the names `lat` and `latitude` case-insensitively in the column names;
- the longitude variable is guessed using `lng`, `long`, `lng`, and `longitude`;

For example, we do not specify the values for the arguments `lat` and `lng` in `addCircles()` below, but the columns `Lat` and `Long` in the data frame `df` will be automatically used:

```{r eval=FALSE}
# add some circles to a map
df = data.frame(Lat = 1:10, Long = rnorm(10))
leaflet(df) %>% addCircles()
# you can also explicitly use Lat and Long
leaflet(df) %>% addCircles(lat = ~ Lat, lng = ~ Long)
```

A map layer may use a different data object to override the data provided in `leaflet()`. We can rewrite the above example as:

```{r eval=FALSE}
leaflet() %>% addCircles(data = df)
# or use df in addCircles() only
leaflet() %>% addCircles(data = df, lat = ~ Lat, lng = ~ Long)
```

Below are examples of using **sp** and **maps**, respectively:

```{r}
library(sp)
Sr1 = Polygon(cbind(c(2, 4, 4, 1, 2), c(2, 3, 5, 4, 2)))
Sr2 = Polygon(cbind(c(5, 4, 2, 5), c(2, 3, 2, 2)))
Sr3 = Polygon(cbind(c(4, 4, 5, 10, 4), c(5, 3, 2, 5, 5)))
Sr4 = Polygon(cbind(c(5, 6, 6, 5, 5), c(4, 4, 3, 3, 4)), hole = TRUE)
Srs1 = Polygons(list(Sr1), "s1")
Srs2 = Polygons(list(Sr2), "s2")
Srs3 = Polygons(list(Sr4, Sr3), "s3/4")
SpP = SpatialPolygons(list(Srs1, Srs2, Srs3), 1:3)
leaflet(height = "300px") %>% addPolygons(data = SpP)
library(maps)
mapStates = map("state", fill = TRUE, plot = FALSE)
leaflet(data = mapStates) %>% addTiles() %>%
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE)
```


## The Formula Interface

The arguments of all layer functions can take normal R objects, such as a numeric vector for the `lat` argument, or a character vector of colors for the `color` argument. They can also take a one-sided formula, in which case the formula will be evaluated using the `data` argument as the environment. For example, `~ x` means the variable `x` in the data object, and you can write arbitrary expressions on the right-hand side, e.g., `~ sqrt(x + 1)`.

```{r eval=FALSE}
m = leaflet() %>% addTiles()
df = data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)
m = leaflet(df) %>% addTiles()
m %>% addCircleMarkers(radius = ~size, color = ~color, fill = FALSE)
m %>% addCircleMarkers(radius = runif(100, 4, 10), color = c('red'))
```

## Shiny

There are two wrapper functions `leafletOutput()` and `renderLeaflet()` in this package for using Leaflet maps in **shiny**.
Compared to the **leaflet** package at <https://github.com/jcheng5/leaflet-shiny>, the **shiny** support in this package is still very limited at the moment, but will be improved in the future.

# UI Layers

The UI elements markers and popups are supported via `addMarkers()` and `addPopups()`, respectively. In the example below, we add a marker with a custom icon, and a popup:

```{r}
leaflet() %>% addTiles() %>%
  addMarkers(174.7690922, -36.8523071, icon = iconList(
    iconUrl = 'http://cran.rstudio.com/Rlogo.jpg',
    iconWidth = 40, iconHeight = 40
  )) %>%
  addPopups(174.7690922, -36.8523071, 'R was born here!')
```

# Raster Layers

Currently tile layers are supported via `addTiles()`. By default, the OpenStreetMap tiles are used, and you can certainly use other tiles of your choice.

A live demo of many tile sets is available [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html). Note that some tile set providers require you to register. See the [project page](https://github.com/leaflet-extras/leaflet-providers) for more information.

```{r}
leaflet() %>%
  addTiles(
    'http://server.arcgisonline.com/ArcGIS/rest/services/World_Topo_Map/MapServer/tile/{z}/{y}/{x}',
    attribution = 'Tiles &copy; Esri &mdash; Esri, DeLorme, NAVTEQ, TomTom, Intermap, iPC, USGS, FAO, NPS, NRCAN, GeoBase, Kadaster NL, Ordnance Survey, Esri Japan, METI, Esri China (Hong Kong), and the GIS User Community') %>%
  setView(-93.65, 42.0285, zoom = 17)
```

# Vector Layers

Vector layers contain elements of vector graphics, such as circles, rectangles, and polylines, etc. All vector layer functions have the arguments `lng` and `lat`, with additional arguments specifying the attributes of the graphical elements. Note the terminology is slightly different with R graphics. For example, the border of elements is controlled by the argument `stroke` (e.g., `leaflet::addRectangles(stroke = FALSE)` is like `graphics::rect(border = NA)`), the `color` argument only specifies the color of the border (in base R, `col`/`color` often means the _fill_ color), and the `weight` argument is like `lwd` in base R graphics.

```{r}
set.seed(123)
m = leaflet() %>% addTiles()
rand_lng = function(n = 10) rnorm(n, -93.65, .01)
rand_lat = function(n = 10) rnorm(n, 42.0285, .01)

# circles (units in metres)
m %>% addCircles(rand_lng(50), rand_lat(50), radius = runif(50, 10, 200))

# circle markers (units in pixels)
m %>% addCircleMarkers(rand_lng(50), rand_lat(50), color = '#ff0000')
m %>% addCircleMarkers(rand_lng(100), rand_lat(100), radius = runif(100, 5, 15))

# rectangles
m %>% addRectangles(
  rand_lng(), rand_lat(), rand_lng(), rand_lat(),
  color = 'red', fill = FALSE, dashArray = '5,5', weight = 3
)

# polylines
m %>% addPolylines(rand_lng(50), rand_lat(50), fill = FALSE)

# polygons
m %>% addPolygons(
  c(rand_lng(3), NA, rand_lng(4), NA, rand_lng(5)),
  c(rand_lat(3), NA, rand_lat(4), NA, rand_lat(5)),
  color = c('red', 'green', 'blue')
)
```

# Other Layers

[GeoJSON](http://geojson.org) is a JSON format for encoding geographic data structures such as points, lines, and polygons. The function `addGeoJSON()` allows us to add a GeoJSON layer to the map, and the layer data is normally represented as an R list. For example, the GeoJSON data

```javascript
var MPoint = {
  "type": "MultiPoint",
  "coordinates": [ [100.0, 0.0], [101.0, 1.0] ]
};
```

can be represented in R as

```{r eval=FALSE}
MPoint = list(
  type = 'MultiPoint',
  coordinates = rbind(c(100.0, 0.0), c(101.0, 1.0))
)
```

You can pass this list to `addGeoJSON()` to draw two points on the map.

For a GeoJSON feature layer, you can use the `style` element in the `properties` list to define its style, and the `popup` element in `properties` to define the popup text. Here is an example:

```{r}
m = leaflet() %>% addCircles(lat = 1:26, lng = 1:26, popup = LETTERS)
shapes = list(
  list(
    type = 'Feature',
    properties = list(
      popup = 'Here are some markers!'
    ),
    geometry = list(
      type = 'MultiPoint',
      coordinates = cbind(10:1, 1:10)
    )
  ),
  list(
    type = 'Feature',
    properties = list(
      style = list(color = 'red', fillColor = 'yellow'),
      popup = 'Here is a polygon, or perhaps a flower...'
    ),
    geometry = list(
      type = 'Polygon',
      coordinates = list(26 + 10 * t(sapply(seq(0, 2 * pi, length = 10), function(x) {
        c(cos(x), sin(x))
      })))
    )
  )
)
m %>% addGeoJSON(shapes)
```

<!--html_preserve--></div><!--/html_preserve-->
