## Awesome Markers

These examples demonstrate the use of `addAwesomeMarkers` to easily add customizable markers, using the [Leaflet.awesome-markers](https://github.com/lvoogdt/Leaflet.awesome-markers) plugin.

### Simple markers

Use `makeAwesomeIcon` with `addAwesomeMarkers` to specify a single icon that should be used for all markers.

```{r}
library(leaflet)

# An icon using the Font Awesome library
icon.fa <- makeAwesomeIcon(icon = "flag", markerColor = "red",
  library="fa", iconColor = "black")

leaflet() %>% addTiles() %>%
  addAwesomeMarkers(
    lng=-118.456554, lat=34.078039,
    label="This is a label",
    icon = icon.fa)

# An icon using the ionicons library
icon.ion <- makeAwesomeIcon(icon = "home", markerColor = "green",
  library="ion")

# Marker + static label using custom label options
leaflet() %>% addTiles() %>%
  addAwesomeMarkers(
    lng=-118.456554, lat=34.078039,
    label="This is a static label",
    labelOptions = labelOptions(noHide = T),
    icon = icon.ion)
```

### Lists of icons with awesomeIcons()

Use `awesomeIcons` instead of `makeAwesomeIcon` to create icons that are identical in most aspects but vary by one or a couple of parameters.

```{r}
cities <- read.csv(textConnection("
City,Lat,Long,Pop
Boston,42.3601,-71.0589,645966
Hartford,41.7627,-72.6743,125017
New York City,40.7127,-74.0059,8406000
Philadelphia,39.9500,-75.1667,1553000
Pittsburgh,40.4397,-79.9764,305841
Providence,41.8236,-71.4222,177994
"))

icon.pop <- awesomeIcons(icon = "users",
  markerColor = ifelse(cities$Pop<500000,"blue","red"),
  library = "fa",
  iconColor = "black"
)

leaflet(cities) %>% addTiles() %>%
  addAwesomeMarkers(lng = ~Long, lat = ~Lat,
    label = ~City,
    icon = icon.pop)
```

### Lists of icons with awesomeIconList()

Finally, you may want to use `awesomeIconList` instead of `awesomeIcons` or `makeAwesomeIcon` if you have two or more very different icons, that you want to select from using a column from your data.

```{r message=FALSE}
# Make a list of icons (from two different icon libraries).
# We'll index into it based on name.
popIcons <- awesomeIconList(
  blue = makeAwesomeIcon(icon="person", library="ion", markerColor="blue"),
  red = makeAwesomeIcon(icon="users", library="fa", markerColor="red")
)

library(dplyr)
cities <- cities %>% mutate(PopCat = ifelse(Pop<500000,"blue","red"))

leaflet(cities) %>% addTiles() %>%
  addAwesomeMarkers(lng = ~Long, lat = ~Lat,
    label = ~City,
    labelOptions = labelOptions(noHide = TRUE),
    icon = ~popIcons[PopCat])
```

