---
title: "Choropleths"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Making choropleths with leaflet is easy. In this example, we'll duplicate the [step-by-step choropleth tutorial](https://leafletjs.com/examples/choropleth/) from the Leaflet.js website.

The final result will look like this (scroll to the end to see the completed code):

```{r, include = FALSE}
library(leaflet)
library(magrittr)
```

```{r fig.height = 4.75, echo = FALSE}
# From http://leafletjs.com/examples/choropleth/us-states.js
states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp")

bins <- c(0, 10, 20, 50, 100, 200, 500, 1000, Inf)
pal <- colorBin("YlOrRd", domain = states$density, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>%g people / mi<sup>2</sup>",
  states$name, states$density
) %>% lapply(htmltools::HTML)

leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN'))) %>%
  addPolygons(
    fillColor = ~pal(density),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>%
  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
    position = "bottomright")
```

### Data source

We'll start by loading the data from JSON. While the Leaflet.js example loads the JSON directly into JavaScript, with the Leaflet R package we instead want to load the data into R.

In this case, we'll use the `geojsonio` package to load the data into `sp` objects, which will let us easily manipulate the geographic features, and their properties, in R.

```{r}
states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp")
class(states)
names(states)
```

As you can see, we now have a `SpatialPolygonsDataFrame` with `name` (state name) and `density` (population density in people/mi<sup>2</sup>) columns from the GeoJSON.

### Basic states map

Next, let's make a basic map with just the outline of the states. For our basemap, we'll use the same `"mapbox.light"` MapBox style that the example does; if you don't have a MapBox account, you can just use `addTiles()` in place of the `addProviderTiles` call, or choose a free provider.

```{r}
m <- leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))
```

We've saved the basic basemap as a separate variable `m` so we can easily iterate on the `addPolygons` call as we work through the rest of the tutorial.

To add uniform polygons with default styling, just call `addPolygons` with no additional arguments.

```{r}
m %>% addPolygons()
```

### Adding some color

Now, let's color the states according to their population density. You have [various options](../colors.html) for mapping data to colors; for this example we'll match the Leaflet.js tutorial by mapping a specific set of bins into RColorBrewer colors.

First, we'll define the bins. This is a numeric vector that defines the boundaries between intervals (`(0,10]`, `(10,20]`, and so on).

Then, we'll call `colorBin` to generate a palette function that maps the RColorBrewer `"YlOrRd"` colors to our bins.

Finally, we'll modify `addPolygons` to use the palette function and the density values to generate a vector of colors for `fillColor`, and also add some other static style properties.

```{r}
bins <- c(0, 10, 20, 50, 100, 200, 500, 1000, Inf)
pal <- colorBin("YlOrRd", domain = states$density, bins = bins)

m %>% addPolygons(
  fillColor = ~pal(density),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7)
```

### Adding interaction

The next thing we'll want is to make the polygons highlight as the mouse passes over them. The `addPolygon` function has a `highlight` argument that makes this simple.

```{r}
m %>% addPolygons(
  fillColor = ~pal(density),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlightOptions = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE))
```

(The Leaflet.js tutorial also adds an event handler that zooms into a state when it's clicked. This isn't currently possible with the Leaflet R package, except with either custom JavaScript or using Shiny, both of which are outside the scope of this example.)

### Custom info

Now let's expose the state names and values to the user.

The Leaflet.js tutorial shows the hovered-over state's information in a custom control. Again, that's possible by adding custom JavaScript or using Shiny, but for this example we'll use the built-in labels feature instead.

We'll generate the labels by handcrafting some HTML, and passing it to `lapply(htmltools::HTML)` so that Leaflet knows to treat each label as HTML instead of as plain text. We'll also set some label options to improve the style of the label element itself.

```{r}
labels <- sprintf(
  "<strong>%s</strong><br/>%g people / mi<sup>2</sup>",
  states$name, states$density
) %>% lapply(htmltools::HTML)

m <- m %>% addPolygons(
  fillColor = ~pal(density),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlightOptions = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
  label = labels,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto"))
m
```

This is the final version of our polygon layer, so let's save the result back to the `m` variable.

### Legend

As our final step, let's add a legend. Because we chose to color our map using `colorBin`, the `addLegend` function makes it particularly easy to add a legend with the correct colors and intervals.

```{r}
m %>% addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
  position = "bottomright")
```

### Complete code

```{r results = 'hide'}
# From http://leafletjs.com/examples/choropleth/us-states.js
states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp")

bins <- c(0, 10, 20, 50, 100, 200, 500, 1000, Inf)
pal <- colorBin("YlOrRd", domain = states$density, bins = bins)

labels <- sprintf(
  "<strong>%s</strong><br/>%g people / mi<sup>2</sup>",
  states$name, states$density
) %>% lapply(htmltools::HTML)

leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN'))) %>%
  addPolygons(
    fillColor = ~pal(density),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) %>%
  addLegend(pal = pal, values = ~density, opacity = 0.7, title = NULL,
    position = "bottomright")
```
