---
title: "Introduction to leaflet"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

[Leaflet](https://leafletjs.com) is one of the most popular open-source JavaScript libraries for interactive maps. It's used by websites ranging from [The New York Times](https://www.nytimes.com/projects/elections/2013/nyc-primary/mayor/map.html) and [The Washington Post](https://www.washingtonpost.com/sf/local/2013/11/09/washington-a-world-apart/) to [GitHub](https://github.blog/2013-06-13-there-s-a-map-for-that/) and [Flickr](https://www.flickr.com/map), as well as GIS specialists like [OpenStreetMap](https://www.openstreetmap.org/), [Mapbox](https://www.mapbox.com/), and [CartoDB](https://carto.com/).

This R package makes it easy to integrate and control Leaflet maps in R.

### Features

* Interactive panning/zooming
* Compose maps using arbitrary combinations of:
    * Map tiles
    * Markers
    * Polygons
    * Lines
    * Popups
    * GeoJSON
* Create maps right from the R console or RStudio
* Embed maps in [knitr](https://yihui.org/knitr/)/[R Markdown](https://rmarkdown.rstudio.com/)/[Quarto](https://quarto.org) documents, and [Shiny](https://shiny.posit.co/) apps
* Easily render spatial objects from the `sp` or `sf` packages, or data frames with latitude/longitude columns
* Use map bounds and mouse events to drive Shiny logic
* Display maps in non spherical mercator projections
* Augment map features using chosen plugins from [leaflet plugins repository](https://leafletjs.com/plugins)

### Installation

To install this R package, run this command at your R prompt:

```{r eval=FALSE}
install.packages("leaflet")
# to install the development version from Github, run
# devtools::install_github("rstudio/leaflet")
```

Once installed, you can use this package at the R console, within [R Markdown](https://rmarkdown.rstudio.com/) documents, and within [Shiny](https://shiny.posit.co/) applications.

### Basic Usage

You create a Leaflet map with these basic steps:

1. Create a map widget by calling `leaflet()`.
2. Add _layers_ (i.e., features) to the map by using layer functions (e.g. `addTiles`, `addMarkers`, `addPolygons`) to modify the map widget.
3. Repeat step 2 as desired.
4. Print the map widget to display it.

Here's a basic example:

```{r}
library(leaflet)

m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
m  # Print the map
```

In case you're not familiar with the pipe ([magrittr](https://github.com/tidyverse/magrittr) or base) operator (`%>%` / `|>`), here is the equivalent without using pipes:

```{r eval=FALSE}
m <- leaflet()
m <- addTiles(m)
m <- addMarkers(m, lng=174.768, lat=-36.852, popup="The birthplace of R")
m
```

### Next Steps

We highly recommend that you proceed to [The Map Widget](map_widget.html) page before exploring the rest of this site, as it describes common idioms we'll use throughout the examples on the other pages.

Although we have tried to provide an R-like interface to Leaflet, you may want to check out the [API documentation](https://leafletjs.com/reference.html) of Leaflet occasionally when the meanings of certain parameters are not clear to you.
