---
title: "Working with projections in Leaflet"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
library(leaflet)
library(sf)
```

**Note:** This functionality is for advanced users and may not be supported across all functions (for example, `addRasterImage()` currently works only with EPSG:3857 Web Mercator).

The Leaflet package expects all point, line, and shape data to be specified in latitude and longitude using WGS 84 (a.k.a. EPSG:4326). By default, when displaying this data it projects everything to [EPSG:3857](https://spatialreference.org/ref/epsg/3857/) and expects that any map tiles are also displayed in EPSG:3857.

For users that need to display data with a different projection, we've integrated the [Proj4Leaflet](https://github.com/kartena/Proj4Leaflet) plugin, which in theory gives Leaflet access to any CRS that is supported by [Proj4js](http://proj4js.org/).

Note that whatever map tiles you use _must_ be designed to work with the CRS of your Leaflet map.

# Defining a custom CRS

Once you've decided on a custom projection, and have map tiles to match it (if necessary), you can use `leafletCRS()` to create a custom projection.

```{r}
crs <- leafletCRS(crsClass = "L.Proj.CRS", code = "ESRI:102003",
  proj4def = "+proj=aea +lat_1=29.5 +lat_2=45.5 +lat_0=37.5 +lon_0=-96 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs",
  resolutions = 1.5^(25:15))
```

The `crsClass` parameter lets you specify the JavaScript constructor to use to generate the Leaflet.js CRS object. In this case, we're using `L.Proj.CRS`, which is the class that comes with Proj4Leaflet. (Only a specific list of CRS classes can be used here; see `?leafletCRS` for more details.)

The `code` parameter is the CRS identifier, usually an EPSG identifier or similar. (For the most part, this doesn't affect us, and you can treat it like documentation; it's primarily used by Proj4Leaflet for purposes that don't apply to the R package.)

The `proj4def` parameter is either a PROJ.4 or WKT string that defines the CRS. If you don't know the PROJ.4 or WKT string, you can generally find these on [epsg.io](https://epsg.io/) or [spatialreference.org](https://spatialreference.org/).

The resulting object can be passed to the `leaflet` function via the parameter `crs = leafletOptions(crs = ...)`.

# Displaying basemap tiles with custom projections

This example shows Gothenberg, Sweden in [EPSG:3006](http://epsg.io/3006) (SWEREF99 TM) projection.

```{r}
epsg3006 <- leafletCRS(crsClass = "L.Proj.CRS", code = "EPSG:3006",
  proj4def = "+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs",
  resolutions = 2^(13:-1), # 8192 down to 0.5
  origin = c(0, 0)
)

tile_url <- "http://api.geosition.com/tile/osm-bright-3006/{z}/{x}/{y}.png"
tile_attrib <- "Map data &copy; <a href='http://www.openstreetmap.org/copyright'>OpenStreetMap contributors</a>, Imagery &copy; 2013 <a href='http://www.kartena.se/'>Kartena</a>"

leaflet(options = leafletOptions(worldCopyJump = FALSE, crs = epsg3006)) %>%
  setView(11.965053, 57.70451, 13) %>%
  addTiles(urlTemplate = tile_url,
    attribution = tile_attrib,
    options = tileOptions(minZoom = 0, maxZoom = 14, continuousWorld = TRUE)) %>%
  addMarkers(11.965053, 57.70451)
```

Again, it's critical to use basemap tiles that are in the same projection that is specified in `leafletCRS()`. In this case the api.geosition.com server does indeed uses EPSG:3006 for its tiles.

We can render a similar map with the default EPSG:3857 projection for comparison. If the custom projection worked correctly, the markers should appear at the same position.

```{r}
leaflet() %>%
  setView(11.965053, 57.70451, 16) %>%
  addTiles() %>%
  addMarkers(11.965053, 57.70451)
```

# Displaying shapes with custom projections

While tiles must be in the same projection as used in `leafletCRS()`, you must always use WGS 84 longitude/latitude data for markers, circles, polygons, and lines. Leaflet will automatically project the coordinates when displaying. 

This example uses data from the `{rnaturalearth}` package and projects it to the [EPSG:9311](https://epsg.io/9311) (US National Atlas Equal Area) coordinate system. Compare the first map, which uses the default CRS, to the second map, which is reprojected.

```{r message=FALSE,warning=FALSE}
north_america <-
  rnaturalearth::countries110 |>
  dplyr::filter(.data$CONTINENT == "North America")

epsg9311 <- leafletCRS(
  crsClass = "L.Proj.CRS",
  code = "EPSG:9311",
  proj4def = "+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +a=6370997 +b=6370997 +units=m +no_defs",
  resolutions = 2 ^ (16:7)
)

pal <- leaflet::colorNumeric(palette = "viridis", domain = north_america$POP_EST)

plot_na_map <- function(opts = leafletOptions()) {
  leaflet(north_america, options = opts) %>%
    addPolygons(
      weight = 1,
      color = "#444444",
      opacity = 1,
      fillColor = ~ pal(POP_EST),
      fillOpacity = 0.7,
      smoothFactor = 0.5,
      label = ~ paste(SOVEREIGNT, POP_EST),
      labelOptions = labelOptions(direction = "auto")
    )
}

plot_na_map()
```

```{r message=FALSE,warning=FALSE}
plot_na_map(opts = leafletOptions(crs = epsg9311))
```


# Polar projections

It's possible to use polar projections, though you may encounter even more problems and incompatibilities with other Leaflet.js plugins than when using other types of custom projections. 

<!-- You can refer to [this set of examples](http://rpubs.com/bhaskarvk/leaflet-polarmaps) by Bhaskar Karambelkar to get started. -->
