---
title: "Raster Images"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Two-dimensional `SpatRaster` objects (from the [`terra` package](https://CRAN.R-project.org/package=terra)) or `RasterLayer` objects (from the [`raster` package](https://CRAN.R-project.org/package=raster)) can be turned into images and added to Leaflet maps using `addRasterImage()`.

`addRasterImage()` works by projecting the `SpatRaster` or `RasterLayer` object to [EPSG:3857](https://spatialreference.org/ref/epsg/3857/) and encoding each cell to an RGBA color, to produce a PNG image. That image is then embedded in the map widget.

It's important that the raster object is tagged with a proper coordinate reference system. Many raster files contain this information, but some do not. Here is how you'd tag a raster object "`r`" which contains WGS84 data:

```r
crs(r) <- "+proj=longlat +datum=WGS84"
```

# Large Raster Warning

Because `addRasterImage()` embeds the image in the map widget, it will increase the size of the generated HTML proportionally. In order to avoid unacceptable download times and memory usage, `addRasterImage()` will error when the PNG is beyond the size indicated by the `maxBytes` argument (defaults to 4 megabytes).

If you have a large raster layer, you can provide a larger number of bytes and see how it goes, or use `terra::resample()`, `raster::resample()`, or `raster::aggregate()` to decrease the number of cells.

# Projection Performance

`addRasterImage()` projects using `terra::project()` or `raster::projectRaster()`, which can take a while on all but the smallest rasters. If you have a large raster layer or expect to call `addRasterImage()` on the same raster layer many times, you can perform the [EPSG:3857](https://spatialreference.org/ref/epsg/3857/) projection yourself (either using `leaflet::projectRasterForLeaflet()` or using another GIS library or program) and call `addRasterImage()` with `project = FALSE`.

Be sure that your pre-projected raster layer is tagged with an accurate extent and CRS, as these values are still needed to place the image in the proper position on the map.

# Coloring

In order to render the raster object as an image, each cell value must be converted to an RGB(A) color. You can specify the color scale using the `colors` argument, which accepts a variety of color specifications:

* The name of a [Color Brewer 2](https://colorbrewer2.org/) palette. If no `colors` argument is provided, then `"Spectral"` is the default.
* A vector that represents the ordered list of colors to map to the data. Any color specification that is accepted by `grDevices::col2rgb()` can be used, including `"#RRGGBB"` and `"#RRGGBBAA"` forms. Example: `colors = c("#E0F3DB", "#A8DDB5", "#43A2CA")`.
* A color scaling function, like those detailed in the [Colors](https://rstudio.github.io/leaflet/articles/colors.html) topic. For example: `colors = colorBin("Greens", domain = NULL, bins = 5, na.color = "transparent")`.

# Example

```{r echo=FALSE}
suppressPackageStartupMessages({
  library(raster)
  r <- suppressWarnings(raster("nc/oisst-sst.nc"))
})
library(leaflet)
# ncdf4 emits partial match warnings
```

```{r eval=FALSE}
library(raster)
library(leaflet)
r <- raster("https://rstudio.github.io/leaflet/nc/oisst-sst.nc")
```

```{r}
pal <- colorNumeric(c("#0C2C84", "#41B6C4", "#FFFFCC"), values(r),
  na.color = "transparent")

leaflet() %>% addTiles() %>%
  addRasterImage(r, colors = pal, opacity = 0.8) %>%
  addLegend(pal = pal, values = values(r),
    title = "Surface temp")
```
