---
pagetitle: Leaflet for R - Using Basemaps
---

## Using Basemaps

Leaflet supports basemaps using [map tiles](https://www.mapbox.com/guides/how-web-maps-work/), popularized by Google Maps and now used by nearly all interactive web maps.

### Default (OpenStreetMap) Tiles

The easiest way to add tiles is by calling `addTiles()` with no arguments; by default, [OpenStreetMap](https://www.openstreetmap.org/) tiles are used.

```{r fig.height=2.5}
m <- leaflet() %>% setView(lng = -71.0589, lat = 42.3601, zoom = 12)
m %>% addTiles()
```

### Third-Party Tiles

Alternatively, many popular free third-party basemaps can be added using the `addProviderTiles()` function, which is implemented using the [leaflet-providers plugin](https://github.com/leaflet-extras/leaflet-providers). See [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html) for the complete set.

As a convenience, leaflet also provides a named list of all the third-party tile providers that are supported by the plugin. This enables you to use auto-completion feature of your favorite R IDE (like RStudio) and not have to remember or look up supported tile providers; just type `providers$` and choose from one of the options. You can also use `names(providers)` to view all of the options.

```{r fig.height=1.25}
m %>% addProviderTiles(providers$Stadia.StamenToner)
m %>% addProviderTiles(providers$CartoDB.Positron)
m %>% addProviderTiles(providers$Esri.NatGeoWorldMap)
```

Note that some tile set providers require you to register; see the [project page](https://github.com/leaflet-extras/leaflet-providers) for more information. You can pass access tokens/keys, and other options, to the tile provider by populating the `options` argument with the `providerTileOptions()` function.

### Custom Tile URL Template

If you happen to have a custom map tile URL template to use, you can provide it as an argument to `addTiles()`.

### WMS Tiles

You can use `addWMSTiles()` to add WMS (Web Map Service) tiles. The map below shows the Base Reflectivity (a measure of the intensity of precipitation occurring) using the WMS from the [Iowa Environmental Mesonet ](http://mesonet.agron.iastate.edu):

```{r fig.height=2.5}
leaflet() %>% addTiles() %>% setView(-93.65, 42.0285, zoom = 4) %>%
  addWMSTiles(
    "http://mesonet.agron.iastate.edu/cgi-bin/wms/nexrad/n0r.cgi",
    layers = "nexrad-n0r-900913",
    options = WMSTileOptions(format = "image/png", transparent = TRUE),
    attribution = "Weather data © 2012 IEM Nexrad"
  )
```

### Combining Tile Layers

You aren't restricted to using a single basemap on a map; you can stack them by adding multiple tile layers. This generally only makes sense if the front tiles consist of semi transparent tiles, or have an adjusted opacity via the `options` argument.

```{r fig.height=1.75}
m %>% addProviderTiles(providers$MtbMap) %>%
  addProviderTiles(providers$Stadia.StamenTonerLines,
    options = providerTileOptions(opacity = 0.35)) %>%
  addProviderTiles(providers$Stadia.StamenTonerLabels)
```
