---
---

## Using Basemaps

Leaflet supports basemaps using [map tiles](https://www.mapbox.com/guides/how-web-maps-work/), popularized by Google Maps and now used by nearly all interactive web maps.

### Default (OpenStreetMap) Tiles

The easiest way to add tiles is by calling `addTiles()` with no arguments; by default, [OpenStreetMap](https://www.openstreetmap.org/) tiles are used.

```{r fig.height=2.5}
m <- leaflet() %>% setView(lng = -71.0589, lat = 42.3601, zoom = 12)
m %>% addTiles() 
```

### Third-Party Tiles

Alternatively, many popular free third-party basemaps can be added using the `addProviderTiles()` function, which is implemented using the [leaflet-providers plugin](https://github.com/leaflet-extras/leaflet-providers). See [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html) for the complete set. Note that some tile set providers require you to register; see the [project page](https://github.com/leaflet-extras/leaflet-providers) for more information.

```{r fig.height=1.25}
m %>% addProviderTiles("Stamen.Toner")
m %>% addProviderTiles("Acetate.terrain")
m %>% addProviderTiles("CartoDB.Positron")
```

### Custom Tile URL Template

If you happen to have a custom map tile URL template to use, you can provide it as an argument to `addTiles()`.

### WMS Tiles [coming soon]

Use `addWMSTiles()` to add WMS tiles, whatever those are. // TODO

### Raster Tiles [coming soon]

You can use Raster objects from the [`raster`](http://cran.r-project.org/web/packages/raster/) package to generate map tile layers on-the-fly. (This only works when Leaflet maps are used with Shiny, as R routines must be invoked as the viewer zooms and pans around the map.)

// TODO: Example

### Combining Tile Layers

You aren't restricted to using a single basemap on a map; you can stack them by adding multiple tile layers. This generally only makes sense if the front tiles consist of semi transparent tiles, or have an adjusted opacity via the `options` argument.

```{r fig.height=1.75}
m %>% addProviderTiles("MtbMap") %>%
  addProviderTiles("Stamen.TonerLines",
    options = providerTileOptions(opacity = 0.35)
  ) %>%
  addProviderTiles("Stamen.TonerLabels")
```
