---
pagetitle: Leaflet for R - Colors
---

## Colors

An important part of spatial visualization is mapping variables to colors. While R has no shortage of built-in functionality to map values to colors, we found that there was enough friction in the process to warrant introducing some wrapper functions that do a lot of the work for you.

To that end, we've created a family of `color*` convenience functions that can be used to easily generate *palette functions*. Essentially, you call the appropriate color function with 1) the colors you want to use and 2) optionally, the range of inputs (i.e. *domain*) that are expected. The color function returns a palette function that can be passed a vector of input values, and it'll return a vector of colors in `#RRGGBB(AA)` format.

```{r}
# Call the color function (colorNumeric) to create a new palette function
pal <- colorNumeric(c("red", "green", "blue"), 1:10)
# Pass the palette function a data vector to get the corresponding colors
pal(c(1,6,9))
```

There are currently three color functions for dealing with continuous input: `colorNumeric`, `colorBin`, and `colorQuantile`; and one for categorical input, `colorFactor`.

### Common parameters

The four color functions all have two required arguments, `palette` and `domain`.

The `palette` argument specifies the colors to map the data to. This argument can take one of several forms:

1. The name of a preset palette from the `RColorBrewer` package, e.g. `"RdYlBu"`, `"Accent"`, or `"Greens"`.
2. The full name of a `viridis` palette: `"viridis"`, `"magma"`, `"inferno"`, or `"plasma"`.
3. A character vector of RGB or named colors, e.g. `palette()`, `c("#000000", "#0000FF", "#FFFFFF")`, `topo.colors(10)`.
4. A function that receives a single value between 0 and 1 and returns a color, e.g.:
`colorRamp(c("#000000", "#FFFFFF"), interpolate="spline")`

The `domain` argument tells the color function the range of input values. You can pass `NULL` here to create a palette function that doesn't have a preset range; the range will be inferred from the data each time you invoke the palette function. If you use a palette function multiple times across different data, it's important to provide a non-`NULL` value for `domain` so the scaling between data and colors is consistent.

### Coloring continuous data

```{r echo=FALSE,results='hide',message=FALSE,warning=FALSE}
library(rgdal)
```
```{r message=FALSE,warning=FALSE,results='hide'}
library(rgdal)

# From http://data.okfn.org/data/datasets/geo-boundaries-world-110m
countries <- readOGR("json/countries.geojson")
map <- leaflet(countries)
```

We've loaded some shape data for countries, including a numeric field `gdp_md_est` which contains GDP estimates.

```{r, fig.height=3.5, fig.width= 4.5, out.width=NULL}
par(mar = c(5,5,0,0), cex = 0.8)
hist(countries$gdp_md_est, breaks = 20, main = "")
```

#### Continuous input, continuous colors (`colorNumeric`)

Let's start by mapping GDP values directly to the `"Blues"` palette from [Color Brewer 2](http://colorbrewer2.org). We'll use the `colorNumeric` function to create a mapping function. The `"Blues"` palette only contains nine colors, but `colorNumeric` interpolates these colors so we get continuous output.

```{r}
# Create a continuous palette function
pal <- colorNumeric(
  palette = "Blues",
  domain = countries$gdp_md_est)
```

The `palette` parameter is the ordered list of colors you will map colors to. In this case we used a Color Brewer palette, but we could've used `c("white", "navy")` or `c("#FFFFFF", "#000080")` for a similar effect. You can also pass more than two colors, for a diverging palette for example. And for maximum flexibility, you can even pass a function that takes a numeric value over the interval [0,1] and returns a color.

The second parameter, `domain`, indicates the set of input values that we are mapping to these colors. For `colorNumeric`, you can provide either a min/max as in this example, or a set of numbers that `colorNumeric` can call `range()` on.

The result is `pal`, a function that can accept numeric vectors with values in the range `range(countries$gdp_md_est)` and return colors in `"#RRGGBB"` format.

```{r}
# Apply the function to provide RGB colors to addPolygons
map %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~pal(gdp_md_est))
```

#### Continuous input, discrete colors (`colorBin` and `colorQuantile`)

`colorBin` maps numeric input data to a fixed number of output colors using binning (slicing the input domain up by value).

You can specify either the exact breaks to use, or the desired number of bins. Note that in the latter case, if `pretty=TRUE` (the default) you'll end up with nice round breaks but not necessarily the number of bins you wanted.

```{r}
binpal <- colorBin("Blues", countries$gdp_md_est, 6, pretty = FALSE)

map %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~binpal(gdp_md_est))
```

`colorQuantile` maps numeric input data to a fixed number of output colors using quantiles (slicing the input domain into subsets with equal numbers of observations).

```{r}
qpal <- colorQuantile("Blues", countries$gdp_md_est, n = 7)
map %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~qpal(gdp_md_est))
```

### Coloring categorical data

For categorical data, use the `colorFactor` function. If the `palette` contains the same number of elements as there are factor levels, then the mapping will be 1:1; otherwise, the palette will be interpolated to produce the desired number of colors.

You can specify the input domain either by passing a factor or character vector to `domain`, or by providing levels directly using the `levels` parameter (in which case the `domain` will be ignored).

```{r}
# Make up some random levels. (TODO: Better example)
countries$category <- factor(sample.int(5L, nrow(countries), TRUE))

factpal <- colorFactor(topo.colors(5), countries$category)

leaflet(countries) %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~factpal(category))
```
