---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(tidy = FALSE)
set.seed(109)
```
[Leaflet](http://leafletjs.com) is an open-source JavaScript library for interactive maps. This R package makes it easy to integrate and control Leaflet maps in R.

## Installation and Use

To install this R package, run this command at your R prompt:

```{r eval=FALSE}
devtools::install_github("rstudio/leaflet")
```

Once installed, you can use this package at the R console, within [R Markdown](http://rmarkdown.rstudio.com/) documents, and within [Shiny](http://shiny.rstudio.com/) applications.

To create a Leaflet map, first you call the function `leaflet()`, then you can add more layers on the map, and manipulate the view. For example,

```{r}
library(leaflet)
m = leaflet() %>% addTiles()
m  # a map with the default OSM tile layer

m = m %>% setView(-93.65, 42.0285, zoom = 17)
m

m %>% addPopups(-93.65, 42.0285, 'Here is the <b>Department of Statistics</b>, ISU')
```

Although we have tried to provide an R-like interface to Leaflet, you may want to check out the [API documentation](http://leafletjs.com/reference.html) of Leaflet occasionally when the meanings of certain parameters are not clear to you.
