---
pagetitle: Leaflet for R - Legends
---

## Legends

The Leaflet package includes convenience functions for creating color legends. In this section, we will build on the example from the [Colors](colors.html) page.

```{r echo=FALSE,results='hide',message=FALSE,warning=FALSE}
library(rgdal)
```
```{r}
library(rgdal)

# From http://data.okfn.org/data/datasets/geo-boundaries-world-110m
countries <- readOGR("json/countries.geojson")
map <- leaflet(countries) %>% addTiles()
```

Use the `addLegend` function to add a legend. The easiest way to use `addLegend` is to provide `pal` (a palette function, as generated from `colorNumeric` et al.) and `values`, and let it calculate the colors and labels for you.

In most cases you will simply be separating the function and argument you passed into `addPolygons(color=...)`, as in this example:

```{r fig.height=2.75}
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = countries$gdp_md_est
)
map %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~pal(gdp_md_est)
  ) %>%
  addLegend("bottomright", pal = pal, values = ~gdp_md_est,
    title = "Est. GDP (2010)",
    labFormat = labelFormat(prefix = "$"),
    opacity = 1
  )
```

The `addLegend()` function is aware of the different types of palette functions, and will create an appropriate default rendering for each type. For example, contrast the legend created for the `colorNumeric`-based palette above with the `colorQuantile`-based palette below. The latter shows probability ranges, with a value range tooltip.

```{r}
qpal <- colorQuantile("RdYlBu", countries$gdp_md_est, n = 5)
map %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.2, fillOpacity = 1,
    color = ~qpal(gdp_md_est)
  ) %>%
  addLegend(pal = qpal, values = ~gdp_md_est, opacity = 1)
```

`addLegend()` has several other parameters that allows you to customize the legend in various ways. Rather than using `pal` and `values`, you can explicitly pass in `colors` and `labels`. You can change the title and color opacity.

You can also conveniently customize the label appearance by passing `labFormat=labelFormat()`. `labelFormat()` has parameters that customize the separator between ranges, the number of digits to render, and prefix/suffix for each label. If your label formatting needs extend beyond what `labelFormat()` can provide, you can also use a custom function as the `labFormat` argument; see the Details section in `?addLegend` for a description.
