---
pagetitle: Leaflet for R - The Map Widget
---

## The Map Widget

The function `leaflet()` returns a Leaflet map widget, which stores a list of objects that can be modified or updated later. Most functions in this package have an argument `map` as their first argument, which makes it easy to use the pipe operator `%>%` in the **magrittr** package, as you have seen from the example in the [Introduction](./).

### Initializing Options
The map widget can be initialized with certain parameters. This is achieved by populating the `options` argument as shown below.

```{r eval=FALSE}
# Set value for the minZoom and maxZoom settings.
leaflet(options = leafletOptions(minZoom = 0, maxZoom = 18))
```

The `leafletOptions()` can be passed any option described in the leaflet [reference document](http://leafletjs.com/reference-1.3.4.html#map-option). Using the `leafletOptions()`, you can set a custom [CRS](https://en.wikipedia.org/wiki/Spatial_reference_system) and have your map displayed in a non spherical mercator projection as described in [projections](projections.html).

### Map Methods

You can manipulate the attributes of the map widget using a series of methods. Please see the help page `?setView` for details.

- `setView()` sets the center of the map view and the zoom level;
- `fitBounds()` fits the view into the rectangle `[lng1, lat1]` -- `[lng2, lat2]`;
- `clearBounds()` clears the bound, so that the view will be automatically determined by the range of latitude/longitude data in the map layers if provided;

### The Data Object

Both `leaflet()` and the map layer functions have an optional `data` parameter that is designed to receive spatial data in one of several forms:

- From base R:
    - lng/lat matrix
    - data frame with lng/lat columns
- From the [**sp** package](http://cran.rstudio.com/package=sp):
    - `SpatialPoints[DataFrame]`
    - `Line`/`Lines`
    - `SpatialLines[DataFrame]`
    - `Polygon`/`Polygons`
    - `SpatialPolygons[DataFrame]`
- From the [**maps** package](http://cran.rstudio.com/package=maps):
    - the data frame from returned from `map()`

The `data` argument is used to derive spatial data for functions that need it; for example, if `data` is a `SpatialPolygonsDataFrame` object, then calling `addPolygons` on that map widget will know to add the polygons from that `SpatialPolygonsDataFrame`.

It is straightforward to derive these variables from **sp** objects since they always represent spatial data in the same way. On the other hand, for a normal matrix or data frame, any numeric column could potentially contain spatial data. So we resort to guessing based on column names:

- the latitude variable is guessed by looking for columns named `lat` or `latitude` (case-insensitive)
- the longitude variable is guessed by looking for `lng`, `long`, or `longitude`

You can always explicitly identify latitude/longitude columns by providing `lng` and `lat` arguments to the layer function.

For example, we do not specify the values for the arguments `lat` and `lng` in `addCircles()` below, but the columns `Lat` and `Long` in the data frame `df` will be automatically used:

```{r eval=FALSE}
# add some circles to a map
df = data.frame(Lat = 1:10, Long = rnorm(10))
leaflet(df) %>% addCircles()
```

You can also explicitly specify the `Lat` and `Long` columns (see below for more info on the `~` syntax):

```{r eval=FALSE}
leaflet(df) %>% addCircles(lng = ~Long, lat = ~Lat)
```

A map layer may use a different data object to override the data provided in `leaflet()`. We can rewrite the above example as:

```{r eval=FALSE}
leaflet() %>% addCircles(data = df)
leaflet() %>% addCircles(data = df, lat = ~ Lat, lng = ~ Long)
```

Below are examples of using **sp** and **maps**, respectively:

```{r message=FALSE}
library(sp)
Sr1 = Polygon(cbind(c(2, 4, 4, 1, 2), c(2, 3, 5, 4, 2)))
Sr2 = Polygon(cbind(c(5, 4, 2, 5), c(2, 3, 2, 2)))
Sr3 = Polygon(cbind(c(4, 4, 5, 10, 4), c(5, 3, 2, 5, 5)))
Sr4 = Polygon(cbind(c(5, 6, 6, 5, 5), c(4, 4, 3, 3, 4)), hole = TRUE)
Srs1 = Polygons(list(Sr1), "s1")
Srs2 = Polygons(list(Sr2), "s2")
Srs3 = Polygons(list(Sr4, Sr3), "s3/4")
SpP = SpatialPolygons(list(Srs1, Srs2, Srs3), 1:3)
leaflet(height = "300px") %>% addPolygons(data = SpP)
library(maps)
mapStates = map("state", fill = TRUE, plot = FALSE)
leaflet(data = mapStates) %>% addTiles() %>%
  addPolygons(fillColor = topo.colors(10, alpha = NULL), stroke = FALSE)
```


### The Formula Interface

The arguments of all layer functions can take normal R objects, such as a numeric vector for the `lat` argument, or a character vector of colors for the `color` argument. They can also take a one-sided formula, in which case the formula will be evaluated using the `data` argument as the environment. For example, `~ x` means the variable `x` in the data object, and you can write arbitrary expressions on the right-hand side, e.g., `~ sqrt(x + 1)`.

```{r eval=FALSE}
m = leaflet() %>% addTiles()
df = data.frame(
  lat = rnorm(100),
  lng = rnorm(100),
  size = runif(100, 5, 20),
  color = sample(colors(), 100)
)
m = leaflet(df) %>% addTiles()
m %>% addCircleMarkers(radius = ~size, color = ~color, fill = FALSE)
m %>% addCircleMarkers(radius = runif(100, 4, 10), color = c('red'))
```
